Decoding the Amplitude-Modulated Part the DCF77 Longwave Time Signal by using the Goertzel Algorithm

### 2024-08-29 paper/main.tex
Started with summary

\section{Demo implementation}
The receiver's structure, shown in Fig. \ref{fig:block_diag_receiver}, was implemented on a STM32F103C8T6, an Arm Cortex-M3 microcontroller, to provide a brief proof of concept.
To for an optical verification, the obtained time code signal is used to turn on and off a Light-Emitting Diode (LED), but this signal
could also be used to decode the DCF77's transmitted time data.
As a tuned antenna together with a pre-amplifier is needed to filter out the DCF77 signal and amplify it to a level where it can be measured by the microcontroller's internal ADC, a minimalistic circuit 
consisting of a tank circuit tuned to \SI{77.5}{\kilo\hertz} and a two-stage transistor amplifier was built based on the circuit shown in \cite{b12}.
The rest of the of the receiver's structure was implemented in software using C and single-precision floating-point for the caluclations.
\par
A timer was set up to create an internal Pulse Width Modulation (PWM) signal to trigger the conversion of the microcontroller's internal ADC at the sampling frequency of \SI{24}{\kilo\hertz}.
The samples are transferred using circular Direct Memory Access (DMA) into a buffer of the size of $480$.
This size was chosen because it is equal to twice the amount of samples that are needed for the analyzation of one segment.
During runtime, the DMA hardware module fills this buffer and two interrupts are executed when the buffer was filled half way and when it is completely full.
Because of the circular setup of the DMA, it fills the buffer continuously and wraps around back to the first address, after reaching the buffer's end.
This allows a continuous operation of the application with a precise sampling frequency while eliminating the need to handle ADC conversions manually. 
As a segment is ready to be analyzed with every interrupt, the operations needed to process the signal processing are executed inside this Interrupt Service Routine (ISR).
The result of each segment after the detector is then used to turn on or off a General Purpose Input Output (GPIO) that is connected to a LED.
\par
Although its rather simplistic implementation, the demo implementation worked right away.
The observed effect of the receiver to adapt to the current receiving conditions by first outputting wrong results because of a
not matching threshold is also noticeable when running the demo application.
After powerup and changes in the environment and therefore during the adaption of the threshold to its new value, the LED shows
arrythmic behavior.
This

The demo implementation worked right away, showing a good sensitivity

@@

\Urlmuskip=0mu plus 1mu
\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: \url{https://de.wikipedia.org/wiki/DCF77}
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/4.42/dcf77.pdf}
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/2009_Bauch_PTBM__DCF77.pdf}
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/5_1988_Hetzel_-_Proc_EFTF_88.pdf}
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: \url{https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html}
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/1407.php}
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: \url{https://ieeexplore.ieee.org/document/6530943}
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: \url{https://en.wikipedia.org/wiki/Goertzel_algorithm}
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: \url{https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf}
\bibitem{b11} R. G. Lyons, ``Setting the 3-dB Cutoff Frequency of an Exponential Averager,'' 2012, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/182.php}
\bibitem{b12} -, ``ATTINY85 als DCF77-Empfänger,'' 2021, Board Entry. [Online]. Available: \url{https://www.mikrocontroller.net/topic/529477}
\end{thebibliography}
\end{document}

>>>>

\section{Demo implementation}
The receiver's structure, shown in Fig. \ref{fig:block_diag_receiver}, was implemented on a STM32F103C8T6, an Arm Cortex-M3 microcontroller, to provide a brief proof of concept.
To for an optical verification, the obtained time code signal is used to turn on and off a Light-Emitting Diode (LED), but this signal
could also be used to decode the DCF77's transmitted time data.
As a tuned antenna together with a pre-amplifier is needed to filter out the DCF77 signal and amplify it to a level where it can be measured by the microcontroller's internal ADC, a minimalistic circuit 
consisting of a tank circuit tuned to \SI{77.5}{\kilo\hertz} and a two-stage transistor amplifier was built based on the circuit shown in \cite{b12}.
The rest of the of the receiver's structure was implemented in software using C and single-precision floating-point for the caluclations.
\par
A timer was set up to create an internal Pulse Width Modulation (PWM) signal to trigger the conversion of the microcontroller's internal ADC at the sampling frequency of \SI{24}{\kilo\hertz}.
The samples are transferred using circular Direct Memory Access (DMA) into a buffer of the size of $480$.
This size was chosen because it is equal to twice the amount of samples that are needed for the analyzation of one segment.
During runtime, the DMA hardware module fills this buffer and two interrupts are executed when the buffer was filled half way and when it is completely full.
Because of the circular setup of the DMA, it fills the buffer continuously and wraps around back to the first address, after reaching the buffer's end.
This allows a continuous operation of the application with a precise sampling frequency while eliminating the need to handle ADC conversions manually. 
As a segment is ready to be analyzed with every interrupt, the operations needed to process the signal processing are executed inside this Interrupt Service Routine (ISR).
The result of each segment after the detector is then used to turn on or off a General Purpose Input Output (GPIO) that is connected to a LED.
\par
Although its rather simplistic implementation, the demo implementation worked right away.
The behavior of the receiver that was observed in the simulation, first outputting a series of wrong results when adapting to the
current receiving conditions, is also noticable when running the demo application.
After powerup or changes in the environment and therefore during the stabilization of the threshold to its new value, the LED shows
arrythmic behavior for approximately \SI{5}{\second} before the application continues to operate correctly.
Other than that, the demo implementation shows that the presented principle also works in reality and that it can be implemented
with low effort.
Based on this structure, the code can be expanded to a full clock by decoding the information from the retrieved time code signal.
The code is available in the paper's GitHub repository \cite{b13}.

\section{Summary}
After the theoretical discussion, the simulation and the proof of concept  presented principle

@@

\Urlmuskip=0mu plus 1mu
\begin{thebibliography}{00}
\bibitem{b1} -, ``DCF77,'' 2024, Wikipedia entry. [Online]. Available: \url{https://de.wikipedia.org/wiki/DCF77}
\bibitem{b2} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77,'' PTB Mitteilungen, 2004, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/4.42/dcf77.pdf}
\bibitem{b3} D. Priester and P. Hetzel and A. Bauch, ``Zeit- und Normalfrequenzverbreitung mit DCF77: 1959-2009 und darüber hinaus,'' PTB Mitteilungen, 2009, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/2009_Bauch_PTBM__DCF77.pdf}
\bibitem{b4} P. Hetzel, ``Time dissemination via the LF transmitter DCF77 using a pseudo-random phase-shift keying of the carrier,'' PTB Braunschweig, 1988, [Online]. Available: \url{https://www.ptb.de/cms/fileadmin/internet/fachabteilungen/abteilung_4/4.4_zeit_und_frequenz/pdf/5_1988_Hetzel_-_Proc_EFTF_88.pdf}
\bibitem{b5} -, ``DCF77 webpage,'' 2024, Official page of the PTB [Online]. Available: \url{https://www.ptb.de/cms/en/ptb/fachabteilungen/abt4/fb-44/ag-442/dissemination-of-legal-time/dcf77.html}
\bibitem{b6} R. G. Lyons, ``Understanding Digital Signal Processing,'' Third Edition, Prentice Hall, 2010.
\bibitem{b9} J. Hoffmann, ``Sampling bandpass signals,'' 2021, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/1407.php}
\bibitem{b10} F. Zaplata and M. Kasal, ``SDR Implementation for DCF77,'', 2013, IEEE Paper. [Online]. Available: \url{https://ieeexplore.ieee.org/document/6530943}
\bibitem{b7} -, ``Goertzel algorithm,'' 2024, Wikipedia entry. [Online]. Available: \url{https://en.wikipedia.org/wiki/Goertzel_algorithm}
\bibitem{b8} K. Banks, ``The Goertzel Algorithm,'' 2002, Internet Article. [Online]. Available: \url{https://courses.cs.washington.edu/courses/cse466/12au/calendar/Goertzel-EETimes.pdf}
\bibitem{b11} R. G. Lyons, ``Setting the 3-dB Cutoff Frequency of an Exponential Averager,'' 2012, Internet Article. [Online]. Available: \url{https://www.dsprelated.com/showarticle/182.php}
\bibitem{b12} -, ``ATTINY85 als DCF77-Empfänger,'' 2021, Board Entry. [Online]. Available: \url{https://www.mikrocontroller.net/topic/529477}
\bibitem{b13} L. Hackmann, ``SiCo1\_Project,'' 2024, GitHub Repository. [Online]. Available: \url{https://github.com/lhckmn/SiCo1_Project}
\end{thebibliography}
\end{document}
